Lexicalized Domain Representations for Neural Machine Translation

### 2019-05-20 2019/emnlp/sparse.tex
resize image, more scores

then optimize each $\theta_i$ (as additional factor used for fine-tuning the model to the domain $D(C_{i})$) the perplexity with respect to true distribution of domain
\begin{equation}
\theta^*_{i} = \displaystyle{\mathop{\argmin}_{\theta_i}}E_{(x,y) \sim D(C_{i})}[-log(p_{\theta_i,\theta^*_s}(y|x,i))]
\end{equation} 
Our first approach to multi-domain machine translation is to build such trivial solution by assigning different regions of word embedding vector to set of domain paramters $\theta_i$ and set of shared parameters $\theta_s$ while keep sharing other parameters between domains.
\subsection{Sparse word embedding}
\label{sec:sparse}
This section explains a new design for word embedding used in multi-domain model. As mentioned above, we separate parameter set into $D+1$ parameter sets correspond to D domains and one shared paramater set. Our first attempt is to split standard vector word embedding into $D+1$ regions named $\theta_i$ for $i \in [1,..,D]$ and $\theta_s$.
\begin{figure}[h]
\center
    \includegraphics[width=\linewidth]{embeddings}
    \caption{Sparse word embeddings. The red region contains units that are activated only when an input from the corresponding domain is presented; the other (green) regions are deactivated; the blue (generic) region is always activated.} 
    \label{fig:network}
  \end{figure}
\fyTodo{Explain vanishing ?}
Figure~\ref{fig:network} presents our proposed design. During backprogation, the \fyTodo{fix notation}{projection matrix $M$} receives gradient 0 at regions corresponding to deactivated regions in the word embedding. Those regions are also masked in forward step, thus do not interfere the training on the domain to which they are not assigned.

@@

\begin{table*}[!h]
\begin{center}
\scalebox{1.0}{
\begin{tabular}{|l|ccc|ccc||c|}
\hline
Model & EMEA & EPPS & ECB & khreshmoi & newstest2014 & iwslt2010 & Avg. \\
\hline
\multicolumn{8}{l}{English$\rightarrow$French} \\
\hline
$Mixed$ & 67.69 $_{47.60}$ & 37.50 & 53.49 & 34.86 & 28.56 & 25.70 & avg \\
\hline
$FT_{EMEA}$ & 76.77 $_{49.43}$ & 17.16 & 11.99 & 29.58 & 11.64 & 11.10 & avg \\
$FT_{EPPS}$ & 20.86 $$ & 37.04 & 24.53 & 31.86 & 28.33 & 11.10 & avg \\
$FT_{ECB}$   & 26.93 & 27.09 & 56.52 & 22.04 & 17.51 & 13.99 & avg \\
\hline
$DC$ & 67.87 $_{45.42}$ & 37.31 & 54.14 & 33.63 & 26.52 & 24.81 & avg\\
\hline
$LDR_{oracle}$   & 74.26 $_{49.90}$ & 37.67 & 54.07 & 33.85 & 28.11 & 25.70 & avg \\
$LDR_{generic}$ & 73.97 $_{49.54}$ & 37.81 & 53.67 & 34.17 & 28.11 & 25.70 & avg \\
$LDR_{rnn}$ & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & avg \\
$LDR_{wrong}$ & 72.95 $_{49.78}$ & 37.62 & 53.35 & khreshmoi & newstest & iwslt & avg\\
\hline
\multicolumn{8}{l}{English$\rightarrow$German} \\
\hline
$Mixed$ & 64.57 $_{42.99}$ & 26.47 & 68.67 & 21.77 & 17.04 & 18.85 & avg \\
\hline
$FT_{EMEA}$ & 68.35 $$ & 17.02 & 32.87 & 21.15 & 11.21 & 13.49 & avg\\
$FT_{EPPS}$ & 36.19 & 26.29 & 40.71 & 19.67 & 16.47 & 19.24 & avg \\
$FT_{ECB}$   & 24.72 & 18.36 & 74.05 & 11.31  & 10.8 & 11.00 & avg \\
\hline
$DC$ & 63.48 $_{42.98}$ & 26.27 & 66.95 & 21.45 & 16.45 & 18.78 & avg \\
\hline
$LDR_{oracle}$ & 70.9 $_{46.12}$ & 26.30 & 68.9 & 21.42 & 16.57 & 18.82 & avg\\
$LDR_{generic}$ & 70.86 $_{45.60}$ & 26.14 & 68.59 & 21.95 & 16.57 & 18.82 & avg\\
$LDR_{rnn}$ & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & avg \\
$LDR_{wrong}$ & 69.51 $_{43.50}$ & 26.31 & 66.86 & khreshmoi & newstest & iwslt & avg \\
\hline
\end{tabular}
} %scalebox
\end{center}
\caption{BPE-detokenized BLEU scores for the Transformer architecture over test sets.}
\end{table*}

\begin{table*}[!h]
\begin{center}
\scalebox{1.0}{
\begin{tabular}{|l|ccc|ccc||c|}
\hline
Model & EMEA & EPPS & ECB & khreshmoi & newstest2014 & iwslt2010 & Avg. \\
\hline
\multicolumn{8}{l}{English$\rightarrow$French} \\
\hline
$Mixed$               & 65.42 $_{45.11}$ & 34.7 & 51.38 & 29.27 & 22.21 & 21.63 & avg \\
\hline
$FT_{EMEA}$     & 72.06 $$ & 18.62 & 16.78 & 25.88 & 10.94 & 5.73 & avg \\
$FT_{EPPS}$      & 35.47 $$ & 34.61 & 39.56 & 27.11 & 21.85 & 20.61 & avg \\
$FT_{ECB}$        & 21.93 $$ & 22.6 & 51.53 & 14.74 & 12.10 & 10.47 & avg \\
\hline
$DC$                   & 68.26 $_{43.76}$ & 35.13 & 50.09 & 27.86 & 20.95 & 21.16 & avg \\
 \hline
$WDCMT$           & 68.76 $_{45.29}$ & 35.71 & 52.75 & 32.03 & 24.63 & 23.28 & avg \\
\hline
$LDR_{oracle}$   & 71.73 $_{46.30}$ & 35.21 & 50.91 & 29.62 & 23.00 & 20.96 & avg\\
$LDR_{generic}$ & 70.59 $_{46.37}$ & 35.22 & 49.68 & 29.51 & 23.00 & 20.96 & avg\\
$LDR_{rnn}$        & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & avg\\
$LDR_{wrong}$   & 62.1 $_{43.29}$ & 34.17 & 48.79 & khreshmoi & newstest & iwslt & avg \\
\hline
\multicolumn{8}{l}{English$\rightarrow$German} \\
\hline
$Mixed$               & 57.37 $_{37.94}$ & 23.1 & 63.54 & 17.39 & 11.51 & 14.29 & avg\\
\hline
$FT_{EMEA}$      & 65.64 $$ & 12.36 & 15.93 & 15.38 & 6.22 & 7.6  & avg\\
$FT_{EPPS}$       & 24.90 $$ & 22.98 & 26.26 & 13.55 & 11.33 & 14.32 & avg \\
$FT_{ECB}$         & 41.8 $$ & 15.97 & 71.07 & 10.77 & 7.66 & 8.07 & avg\\
\hline
$DC$                    & 62.53 $_{39.25}$ & 23.74 & 65.71 & 16.88 & 11.15 & 14.97 & avg\\
\hline
WDCMT               & 64.05 $_{38.54}$ & 22.82 & 67.00 & 15.94 & 11.45 & 13.66 & avg\\
\hline
$LDR_{oracle}$   & 63.43 $_{40.04}$ & 22.66 & 64.40 & 16.99 & 11.45 & 14.26 & avg\\
$LDR_{generic}$ & 63.27 $_{39.75}$ & 22.62 & 63.70 & 17.26 & 11.45 & 14.26 & avg\\
$LDR_{rnn}$        & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & avg\\
$LDR_{wrong}$   & 56.84 $_{37.05}$ & 22.06 & 61.66 & khreshmoi & newstest & iwslt & avg \\
\hline
\end{tabular}
} %scalebox
\end{center}
\caption{BPE-detokenized BLEU scores for the RNN architecture over test sets.}
\end{table*}

@@

\begin{table*}[!h]
\begin{center}
\scalebox{1.0}{
\begin{tabular}{|l|ccc|ccc|}
\hline
$LDR$ & EMEA & EPPS & ECB & khreshmoi & newstest2014 & iwslt2010 \\
\hline
\multicolumn{7}{l}{English$\rightarrow$French} \\
\hline
 size=4   & 74.65 $_{49.61}$ & 37.42 & 54.49 & 33.68 & 28.03 & 25.28 \\
 size=8   & 74.26 $_{49.90}$ & 37.67 & 54.07 & 33.85 & 28.11 & 25.70 \\
 size=16 & 74.15 $_{49.10}$ & 37.78 & 54.56 & 33.40 & 28.10 & 25.33 \\
 size=32 & 75.10 $_{48.61}$ & 37.64 & 54.29 & 33.20 & 27.99 & 25.44 \\
\hline
\end{tabular}
} %scalebox
\end{center}
\caption{BPE-detokenized BLEU scores for the RNN architecture over test sets.}
\end{table*}

@@

Multi-domain Machine Translation has largely interested NLP community by its promising applications in industry and its relation to fundamental problems in theory. Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{;.\\ Data centric category attempts to collect related-domain data from large existing in-domain corpora. The methods in this category can be listed as following:
\fyTodo{Organize refs}
\begin{itemize}
\item score relatedness of sentences to the target domain by Language Model \cite{Moore2010selection,Axelrod2011domain,Duh2013selection}.
\item generating pseudo parallel data \cite{Utiyama2003measure,Wang2016connecting,Wang2014neural}
\end{itemize}
On the other hand, model-centric approaches focus on NMT models that are specialized for domain adaptation. The novelty can be among following features
\begin{itemize}
\item Trainining objective \cite{Luong2015SNMT,Senrich2016Mono,Wang2017Instance,Chen2017Cost,Miceli2017Regularize,Zhang18sentence}
\item Architecture \cite{gulcehre2016monolingual,Zhang16topicinformed,Kobus17domaincontrol,Britz2017mixing,Biao2017CARENMT,Britz2017mixing,Thompson18freezing,Michel2018extreme}.
\item The decoding algorithm \cite{gulcehre2016monolingual,Khayrallah2017lattice}
\end{itemize}

Beside these works\fyTdodo{Works= oeuvres d'art}, we could also consider works in other tasks such as sequence labeling tasks \cite{Daume07frustratingly}; learning multiple visual domains \cite{Rebuffi2017Visual}.

>>>>

then optimize each $\theta_i$ (as additional factor used for fine-tuning the model to the domain $D(C_{i})$) the perplexity with respect to true distribution of domain
\begin{equation}
\theta^*_{i} = \displaystyle{\mathop{\argmin}_{\theta_i}}E_{(x,y) \sim D(C_{i})}[-log(p_{\theta_i,\theta^*_s}(y|x,i))]
\end{equation} 
Our first approach to multi-domain machine translation is to build such trivial solution by assigning different regions of word embedding vector to set of domain paramters $\theta_i$ and set of shared parameters $\theta_s$ while keep sharing other parameters between domains.
\subsection{Sparse word embedding}
\label{sec:sparse}
This section explains a new design for word embedding used in multi-domain model. As mentioned above, we separate parameter set into $D+1$ parameter sets correspond to D domains and one shared paramater set. Our first attempt is to split standard vector word embedding into $D+1$ regions named $\theta_i$ for $i \in [1,..,D]$ and $\theta_s$.
\begin{figure}
\center
    \includegraphics[width=0.5\linewidth]{embeddings}
    \caption{Sparse word embeddings. The red region contains units that are activated only when an input from the corresponding domain is presented; the other (green) regions are deactivated; the blue (generic) region is always activated.} 
    \label{fig:network}
  \end{figure}
\fyTodo{Explain vanishing ?}
Figure~\ref{fig:network} presents our proposed design. During backprogation, the \fyTodo{fix notation}{projection matrix $M$} receives gradient 0 at regions corresponding to deactivated regions in the word embedding. Those regions are also masked in forward step, thus do not interfere the training on the domain to which they are not assigned.

@@

\begin{table*}[!h]
\begin{center}
\scalebox{1.0}{
\begin{tabular}{|l|ccc|ccc||c|}
\hline
Model & EMEA & EPPS & ECB & khreshmoi & newstest2014 & iwslt2010 & Avg. \\
\hline
\multicolumn{8}{l}{English$\rightarrow$French} \\
\hline
$Mixed$ & 67.69$_{47.60}$ & 37.50 & 53.49 & 34.86 & 28.56 & 25.70 & avg \\
\hline
$FT_{EMEA}$ & 76.77$_{49.43}$ & 17.16 & 11.99 & 29.58 & 11.64 & 11.10 & avg \\
$FT_{EPPS}$ & 20.86 & 37.04 & 24.53 & 31.86 & 28.33 & 11.10 & avg \\
$FT_{ECB}$   & 26.93 & 27.09 & 56.52 & 22.04 & 17.51 & 13.99 & avg \\
\hline
$DC$ & 67.87$_{45.42}$ & 37.31 & 54.14 & 33.63 & 26.52 & 24.81 & avg\\
\hline
$LDR_{oracle}$   & 74.26$_{49.90}$ & 37.67 & 54.07 & 33.85 & 28.11 & 25.70 & avg \\
$LDR^{v2}_{oracle}$   & 74.95$_{49.38}$ & 37.35 & 55.91 & 33.57 & 27.59 & 25.27 & avg \\
$LDR_{generic}$ & 73.97$_{49.54}$ & 37.81 & 53.67 & 34.17 & 28.11 & 25.70 & avg \\
$LDR_{rnn}$ & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & avg \\
$LDR_{wrong}$ & 72.95$_{49.78}$ & 37.62 & 53.35 & khreshmoi & newstest & iwslt & avg\\
\hline
\multicolumn{8}{l}{English $\rightarrow$German} \\
\hline
$Mixed$ & 64.57$_{42.99}$ & 26.47 & 68.67 & 21.77 & 17.04 & 18.85 & avg \\
\hline
$FT_{EMEA}$ & 68.35 & 17.02 & 32.87 & 21.15 & 11.21 & 13.49 & avg\\
$FT_{EPPS}$ & 36.19 & 26.29 & 40.71 & 19.67 & 16.47 & 19.24 & avg \\
$FT_{ECB}$   & 24.72 & 18.36 & 74.05 & 11.31  & 10.8 & 11.00 & avg \\
\hline
$DC$ & 63.48$_{42.98}$ & 26.27 & 66.95 & 21.45 & 16.45 & 18.78 & avg \\
\hline
$LDR_{oracle}$ & 70.9$_{46.12}$ & 26.30 & 68.9 & 21.42 & 16.57 & 18.82 & avg \\
$LDR^{v2}_{oracle}$ & 71.31$_{45.23}$ & 25.98 & 73.74 & 21.6 & 16.37 & 19.2 & avg \\
$LDR_{generic}$ & 70.86$_{45.60}$ & 26.14 & 68.59 & 21.95 & 16.57 & 18.82 & avg\\
$LDR_{rnn}$ & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & avg \\
$LDR_{wrong}$ & 69.51$_{43.50}$ & 26.31 & 66.86 & khreshmoi & newstest & iwslt & avg \\
\hline
\end{tabular}
} %scalebox
\end{center}
\caption{BPE-detokenized BLEU scores for the Transformer architecture over test sets.}
\end{table*}

\begin{table*}[!h]
\begin{center}
\scalebox{1.0}{
\begin{tabular}{|l|ccc|ccc||c|}
\hline
Model & EMEA & EPPS & ECB & khreshmoi & newstest2014 & iwslt2010 & Avg. \\
\hline
\multicolumn{8}{l}{English$\rightarrow$French} \\
\hline
$Mixed$               & 65.42$_{45.11}$ & 34.7 & 51.38 & 29.27 & 22.21 & 21.63 & avg \\
\hline
$FT_{EMEA}$     & 72.06 $ $ & 18.62 & 16.78 & 25.88 & 10.94 & 5.73 & avg \\
$FT_{EPPS}$      & 35.47 $ $ & 34.61 & 39.56 & 27.11 & 21.85 & 20.61 & avg \\
$FT_{ECB}$        & 21.93 $ $ & 22.6 & 51.53 & 14.74 & 12.10 & 10.47 & avg \\
\hline
$DC$                   & 68.26 $_{43.76}$ & 35.13 & 50.09 & 27.86 & 20.95 & 21.16 & avg \\
 \hline
$WDCMT$           & 68.76 $_{45.29}$ & 35.71 & 52.75 & 32.03 & 24.63 & 23.28 & avg \\
\hline
$LDR_{oracle}$   & 71.73 $_{46.30}$ & 35.21 & 50.91 & 29.62 & 23.00 & 20.96 & avg\\
$LDR_{generic}$ & 70.59 $_{46.37}$ & 35.22 & 49.68 & 29.51 & 23.00 & 20.96 & avg\\
$LDR_{rnn}$        & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & avg\\
$LDR_{wrong}$   & 62.1 $_{43.29}$ & 34.17 & 48.79 & khreshmoi & newstest & iwslt & avg \\
\hline
\multicolumn{8}{l}{English$\rightarrow$German} \\
\hline
$Mixed$               & 57.37 $_{37.94}$ & 23.1 & 63.54 & 17.39 & 11.51 & 14.29 & avg\\
\hline
$FT_{EMEA}$      & 65.64 $ $ & 12.36 & 15.93 & 15.38 & 6.22 & 7.6  & avg\\
$FT_{EPPS}$       & 24.90 $ $ & 22.98 & 26.26 & 13.55 & 11.33 & 14.32 & avg \\
$FT_{ECB}$         & 41.8 $ $ & 15.97 & 71.07 & 10.77 & 7.66 & 8.07 & avg\\
\hline
$DC$                    & 62.53 $_{39.25}$ & 23.74 & 65.71 & 16.88 & 11.15 & 14.97 & avg\\
\hline
WDCMT               & 64.05 $_{38.54}$ & 22.82 & 67.00 & 15.94 & 11.45 & 13.66 & avg\\
\hline
$LDR_{oracle}$   & 63.43 $_{40.04}$ & 22.66 & 64.40 & 16.99 & 11.45 & 14.26 & avg\\
$LDR_{generic}$ & 63.27 $_{39.75}$ & 22.62 & 63.70 & 17.26 & 11.45 & 14.26 & avg\\
$LDR_{rnn}$        & EMEA & EPPS & ECB & khreshmoi & newstest & iwslt & avg\\
$LDR_{wrong}$   & 56.84 $_{37.05}$ & 22.06 & 61.66 & khreshmoi & newstest & iwslt & avg \\
\hline
\end{tabular}
} %scalebox
\end{center}
\caption{BPE-detokenized BLEU scores for the RNN architecture over test sets.}
\end{table*}

@@

\begin{table*}[!h]
\begin{center}
\scalebox{1.0}{
\begin{tabular}{|l|ccc|ccc|}
\hline
$LDR$ & EMEA & EPPS & ECB & khreshmoi & newstest2014 & iwslt2010 \\
\hline
\multicolumn{7}{l}{English$\rightarrow$French} \\
\hline
 size=4   & 74.65 $_{49.61}$ & 37.42 & 54.49 & 33.68 & 28.03 & 25.28 \\
 size=8   & 74.26 $_{49.90}$ & 37.67 & 54.07 & 33.85 & 28.11 & 25.70 \\
 size=16 & 74.15 $_{49.10}$ & 37.78 & 54.56 & 33.40 & 28.10 & 25.33 \\
 size=32 & 75.10 $_{48.61}$ & 37.64 & 54.29 & 33.20 & 27.99 & 25.44 \\
 size=64 & 74.5 $_{50.17}$ & 37.27 & 54.5 & 33.26 & 27.95 & 25.65 \\
\hline
\end{tabular}
} %scalebox
\end{center}
\caption{BPE-detokenized BLEU scores for the RNN architecture over test sets.}
\end{table*}

@@

Multi-domain Machine Translation has largely interested NLP community by its promising applications in industry and its relation to fundamental problems in theory. Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{Chu18asurvey}.\\ Data centric category attempts to collect related-domain data from large existing in-domain corpora. The methods in this category can be listed as following:
\fyTodo{Organize refs}

\begin{itemize}
\item score relatedness of sentences to the target domain by Language Model \cite{Moore2010selection,Axelrod2011domain,Duh2013selection}.
\item generating pseudo parallel data \cite{Utiyama2003measure,Wang2016connecting,Wang2014neural}
\end{itemize}
On the other hand, model-centric approaches focus on NMT models that are specialized for domain adaptation. The novelty can be among following features

\begin{itemize}
\item Trainining objective\cite{Luong2015SNMT,Senrich2016Mono,Wang2017Instance,Chen2017Cost,Miceli2017Regularize,Zhang18sentence}
\item Architecture %\cite{gulcehre2016monolingual,Zhang16topicinformed,Kobus17domaincontrol,Britz2017mixing,Biao2017CARENMT,Britz2017mixing,Thompson18freezing,Michel2018extreme}.
\item The decoding algorithm %\cite{gulcehre2016monolingual,Khayrallah2017lattice}
\end{itemize}

%Beside these works \fyTdodo{Works= oeuvres d'art}, we could also consider works in other tasks such as sequence labeling tasks \cite{Daume07frustratingly}; learning multiple visual domains \cite{Rebuffi2017Visual}.
